---
output:
  pdf_document: default
  html_document: default
---
# Final Project #

# Tianqi Luo #

# Predict diastolic blood pressure with MCMC model #



## Introduction ##

**In the United States, cardiovascular disease, or heart disease, has been the leading cause of death for both men and women for decades. One of the indicators of the cardiovascular disease is the diastolic blood pressure, the pressure in the arteries when heart rests between beats. For this reason, it's important to look at the diastolic blood pressure as an important indicator for cardiovascular diseases. In this project, I'm trying to build a Bayesian MCMC model that accurately predicts the diastolic blood pressure with several variables by using the cardiovascular dataset provided in Kaggle.**




## Data ##

**In my final project, I'm looking into the cardiovascular disease dataset("https://www.kaggle.com/sulianova/cardiovascular-disease-dataset") on Kaggle. The dataset consists of 70000 records of patients data. It takes into account of their age, height, weight, gender, and measures their diastolic blood pressures, cholesterol levels, and glucose levels.**


## Purpose ##

**To construct an MCMC model that can accurately demonstrate the relationship between the diastolic blood pressure and other variables.**





## Variables Introduced In the Model ##

**height: Patient's height**

**weight: Patient's weight**

**gender: Patient's gender**

**ap_lo: Diastolic blood pressure**

**Glucose: Patient's glucose level**

**Cholesterol: Patient's cholesterol level**







## Analysis ##

### Run the BUGS code in R ###

```{r}
library(R2WinBUGS)
cardio = read.csv("./cardio.csv", sep = ";")
```

**Subset the cardio dataset**
```{r}
cardio = cardio[1:1000, ]
```

**Initiate the variables**
```{r}
gender = cardio$gender
x1 = cardio$weight
x2 = cardio$height
n = length(gender)
y = cardio$ap_lo
J = length(unique(gender))
```


**Write a varying-intercept(by gender) model for diastolic blood pressure using weight and height as predictors**
```{r}
model = function() {
 for (i in 1:n){
     y[i] ~ dnorm (y.hat[i], tau.y)
     y.hat[i] <- a[gender[i]] + b*x1[i] + c*x2[i]
}
b ~ dnorm (0, .0001) 
c ~ dnorm (0, .0001) 
tau.y <- pow(sigma.y, -2) 
sigma.y ~ dunif (0, 100) 

for (j in 1:J){
   a[j] ~ dnorm (mu.a, tau.a) 
  }
   mu.a ~ dnorm (0, .0001) 
   tau.a <- pow(sigma.a, -2) 
   sigma.a ~ dunif (0, 100)
}

write.model(model, "cardio.1.bug")
model.file1 = paste(getwd(),"cardio.1.bug", sep="/")
file.show("cardio.1.bug")
```


**Initiate the parameters**
```{r}
cardio.1.data <- list ("n", "J", "y", "gender", "x1", "x2") 
cardio.1.inits <- function (){
list (a=rnorm(J), b=rnorm(1), c=rnorm(1), mu.a=rnorm(1), sigma.y=runif(1), sigma.a=runif(1))}
cardio.1.parameters <- c ("a", "b", "c", "sigma.y", "sigma.a")
```

**Run the BUGS code in WinBUGS**
```{r}
set.seed(4744469)
cardio.1 <- bugs (cardio.1.data, cardio.1.inits, cardio.1.parameters, "cardio.1.bug", n.chains=3, n.iter=10000, bugs.directory = "/Users/tianqiluo49/winbugs14_full_patched-3/WinBUGS14/", debug=TRUE)
```

**Print the results of the va®¥ing-intercept varying-slope model**
```{r}
print(cardio.1)
```

**Since the effect size for b is large and positive, and for c is small and negative, we can conclude that the diastolic blood pressure has a strong positive relationship with weight, but a weak negative relationship with height**

**Plot the results**
```{r}
plot(cardio.1)
```

**We can see from the graphs that, the chains all mix very well and the 80% intervals explain the parameters really well as well, so this is a great model to use**




**Write the BUGS code for varying-intercept with weight, height, cholesterol level and glucose level, and use age as a group-level predictor**
```{r}
model = function() {
 for (i in 1:n){
     y[i] ~ dnorm (y.hat[i], tau.y)
     y.hat[i] <- a[gender[i]] + b*x1[i] + c*x2[i] + d*x3[i] + e*x4[i]
}
b ~ dnorm (0, .0001) 
c ~ dnorm (0, .0001) 
d ~ dnorm (0, .0001) 
e ~ dnorm (0, .0001) 
tau.y <- pow(sigma.y, -2) 
sigma.y ~ dunif (0, 100) 

for (j in 1:J){
   a[j] ~ dnorm (mu.a, tau.a) 
   a.hat[j] <- g.0 + g.1*u[j]
}

   g.0 ~ dnorm (0, .0001) 
   g.1 ~ dnorm (0, .0001) 
   

   mu.a ~ dnorm (0, .0001) 
   tau.a <- pow(sigma.a, -2) 
   sigma.a ~ dunif (0, 100)
}

write.model(model, "cardio.2.bug")
model.file1 = paste(getwd(),"cardio.2.bug", sep="/")
file.show("cardio.2.bug")
```

**Initiate the variables cholesterol and gluc**
```{r}
x3 = cardio$cholesterol
x4 = cardio$gluc
```

**Greate a group variable for age based on gender**
```{r}
u = cardio[unique(cardio$gender), "age" ]
```

**Initiate the parameters**
```{r}
cardio.2.data <- list ("n", "J", "y", "gender", "x1", "x2", "u", "x3", "x4") 
cardio.2.inits <- function (){
list (a=rnorm(J), b=rnorm(1), c=rnorm(1), d = rnorm(1), e = rnorm(1), g.0=rnorm(1), g.1=rnorm(1), mu.a=rnorm(1), sigma.y=runif(1), sigma.a=runif(1))}
cardio.2.parameters <- c ("a", "b", "c", "d", "e", "sigma.y", "sigma.a")
```


**Run the BUGS code in WinBUGS**
```{r}
set.seed(4744469)
cardio.2 <- bugs (cardio.2.data, cardio.2.inits, cardio.2.parameters, "cardio.2.bug", n.chains=3, n.iter=10000, bugs.directory = "/Users/tianqiluo49/winbugs14_full_patched-3/WinBUGS14/", debug=TRUE)
```

**Print out the results**
```{r}
print(cardio.2)
```

**We can see from this model that the effect sizes for b, d, e are all very large, so we can see that diastolic blood pressure are strongly correlated with weight, glucose level and cholesterol level. On average, the diastolic blood pressure has a moderate positive relationship with weight, weak positive relationship with height, strong positive relationship with cholesterol level, and strong negative relationship with glucose level**

**Compared to model 1, the second model has generally bigger effect sizes for the variables, while having similar DIC and standard deviations**

**Plot the results**
```{r}
plot(cardio.2)
```

**We can see that not only do the chains mix better, but the 80% interval of the parameters are also smaller and more precise.**




## Conclusion From the Project ##

**1. From the above models, we can conclude that weight, glucose level, and cholesterol levels are all strong predictors of diastolic blood pressure. While weight and cholesterol have positive relationships with diastolic blood pressure, glucose level has negative.**

**2. Model 2 is significantly better than model 1 because it includes more variables with large effect sizes, also the general effect sizes are bigger across all variables.**
