---
output:
  pdf_document: default
  html_document: default
---

## Stat 618 ## 

## Tianqi Luo ##

## HW1 ##

## 2 ##

**(1)**

```{r}
library(tidyverse)
girls = read.table("./girls.txt")
head(girls)
```


```{r}
girls %>% 
  select(V2) %>%
  rename(girlbirths = 1) ->
  tidy_girls

head(tidy_girls)
```


```{r}
actual_sd = sd(tidy_girls$girlbirths)
p = mean(tidy_girls$girlbirths)
actual_sd
p
```

```{r}
n = 3900
expected_sd = sqrt((p * (1-p))/n)
expected_sd
```

**The actual standard deviation is 0.00641, the theoretical standard deviation is 0.00800**



**(2)**

```{r}
alpha = 0.05
df = 23
chi_square_0.025 = qchisq(1 - alpha/2, df)
chi_square_0.975 = qchisq(alpha/2, df)

chi_square_0.025
chi_square_0.975
```

```{r}
lower_bound = sqrt((df*(expected_sd)^2)/(chi_square_0.025))
lower_bound 
```

```{r}
upper_bound = sqrt((df*(expected_sd)^2)/(chi_square_0.975))
upper_bound
```

**We can see that the 95% CI for the theoretical standard deviation is (0.00622, 0.0112), the actual standard deviation is 0.00641. Since the actual standard deviation falls within the 95% CI for the theoretical standard deviation, the difference is not statistically significant**


## 3 ##

```{r}
n1 = 1000
N1 = 20
```


```{r}
mean_single = (0.5) * (0 + 1)
var_single = (1/12) * (1-0)^2

total_mean = mean_single * N1 
total_sd = sqrt(var_single * N1)

total_mean
total_sd
```

```{r}
set.seed(4744470)
sim = runif(N1 * n1)
mat = matrix(sim, nrow = n1, ncol = N1)
```

```{r}
data = data.frame(rowSums(mat))

data %>%
  rename(Sum = 1) ->
  tidy_data
```

```{r}
ggplot(tidy_data, aes(x = Sum)) + 
  geom_histogram(aes(y = ..density..), binwidth = 1, alpha = 0.1) + 
  geom_density(lty = 2) + 
  stat_function(fun=dnorm, args=list(mean=total_mean, sd=total_sd), lwd = 2, col = 'blue', alpha = 0.5) + 
  xlab("Sum X") + 
  ylab("Density")
```

**The distribution of the curve and the distribution of the histogram for the simulation are very similar. However, it does vary quite a bit at the mean. With the increase number of simulations, the histogram will become narrower and more similar with the curve. When n is big enough, the histogram will completely overlap with the curve**


## 4 ##

```{r}
N2 = 100
n2= 1000
men_mean = 69.1
women_mean = 63.7
men_sd = 2.9
women_sd = 2.7
```

```{r}
set.seed(4744469)
sim_men = rnorm(n2 * N2, mean = men_mean, sd = men_sd)
sim_women = rnorm(n2 * N2, mean = women_mean, sd = women_sd)
mat_men = matrix(sim_men, nrow = n2, ncol = N2)
mat_women = matrix(sim_women, nrow = n2, ncol = N2)
```

```{r}
height_data = data.frame(men_height_mean = rowMeans(mat_men), women_height_mean = rowMeans(mat_women))

head(height_data)
```

```{r}
height_data %>%
  mutate(height_difference = men_height_mean - women_height_mean) ->
  final_height_data
```

```{r}
ggplot(final_height_data, aes(x = height_difference)) + 
  geom_histogram(binwidth = 0.1) + 
  xlab("Height Difference") + 
  ylab("Count")
```

**Calculate the observed mean and observed standard deviation for the height difference**
```{r}
final_height_data %>%
  summarize(mean_height_difference = mean(height_difference), sd_height_difference = sd(height_difference)) ->
  simulation_values

simulation_values
```

**We can see the observed mean of the height difference is 5.384, and the observed standard deviation of height difference is 0.403** 


**Theoretical mean and sd**
```{r}
theoretical_mean = men_mean - women_mean
theoretical_sd = sqrt((men_sd)^2 + (women_sd)^2)
data.frame(c(theoretical_mean, theoretical_sd))
```

**We can see the theoretical mean of height difference is 5.400, and the theoretical standard deviation of height difference is 3.962**

**So, we can see that the theoretical and observed values for mean and standard deviation of height differences are extremely similar**





  