---
output:
  pdf_document: default
  html_document: default
---


## Stat 618 HW3 ##

## Tianqi Luo ##


### Problem 11.4 ###

```{r}
library(tidyverse)
library(lubridate)
library(covBM)
library(lme4)
library(arm)
library(ggplot2)
```

### (a) ###

**Load the cd4 data from the covBM package**
```{r}
cd4_scaled = covBM::cd4
```

```{r}
head(cd4_scaled)
```


```{r}
fit_cd4 = lm(sqrtcd4 ~ t, cd4_scaled)
display(fit_cd4)
```

**Plot the variables with sqrtcd4 with t, plot newpid in different colors**
```{r}
p = ggplot(cd4_scaled, aes(x = t, y = sqrtcd4, color = as.factor(newpid))) + 
  geom_point() + 
  xlab("Time") + 
  ylab("Square Root cd4") + 
  theme(legend.position = "none")

p
```

### (b) ###

**Glimpse through the coefficients of every single linear regression function grouped by newpid**
```{r}
estimated_values = summarize(group_by(cd4_scaled, as.factor(newpid)), intercept = lm(sqrtcd4 ~ t)$coeff[1], slope = lm(sqrtcd4 ~ t)$coeff[2], fit_values = mean(lm(sqrtcd4 ~ t)$fitted.values))

estimated_values %>%
  rename(newpid = 1) -> estimated_values

head(estimated_values)
```

**Plot the lines for the graph**
```{r}
p + geom_smooth(method = "lm", formula = y~x, se = F)
```



### (c) ###

**Mutate the treatment variable as a factor, and create a data table. Use the summarize function to summarize the coefficients and fitted values of the linear values by treatment groups**


```{r}
library(data.table)
dat = data.table(x = cd4_scaled$baseage, y = cd4_scaled$sqrtcd4, treatment = as.factor(cd4_scaled$treatmnt))
```


```{r}
first_step = summarize(group_by(dat, treatment), intercept = lm(y~x)$coeff[1], slope = lm(y~x)$coeff[2], fit_values = mean(lm(y~x)$fitted.values))

head(first_step)
```

**Treat the coefficients obtained by the first step as data, and use the y_average grouped by treatment as response variable, fit the model**

```{r}
second_step = lm(fit_values ~ intercept + slope, first_step)
display(second_step)
```

**We can see that the two-step process is very flawed as it doesn't take into account of the standard error within the grouping variable**




## Problem 12.2 ##

### (a) ###

```{r}
cd4 = read_csv("./allvar.csv")

head(cd4)
```

```{r}
cd4$VDATE = mdy(cd4$VDATE)
fit_1 = lmer(CD4PCT ~ VDATE + (1 | newpid), data = subset(cd4, !is.na(treatmnt) & !is.na(baseage)))
display(fit_1)
```


### (b) ###

```{r}
fit_2 = lmer(CD4PCT ~ VDATE + treatmnt + baseage + (1 | newpid), data = cd4)

display(fit_2)
```

### (c) ###

**Extract the intercepts from fit_1**
```{r}
df1 = coef(fit_1)$newpid[, 1, drop = F]

head(df1)
```

**Extract the intercepts for fit_2**
```{r}
df2 = coef(fit_2)$newpid[, 1, drop = F]

head(df2)
```

**Change the names of the dataframe**
```{r}
names(df1) = "intercept"
names(df2) = "intercept"
```

**Calculate the upper CI and lower CI of model 1**
```{r}
df1$lower_ci = df1$intercept - (1.96 * se.ranef(fit_1)$newpid[,1])
df1$upper_ci = df1$intercept + (1.96 * se.ranef(fit_1)$newpid[,1])
```

**Calculate the upper CI and lower CI of model 2**
```{r}
df2$lower_ci = df2$intercept - (1.96 * se.ranef(fit_2)$newpid[,1])
df2$upper_ci = df2$intercept + (1.96 * se.ranef(fit_2)$newpid[,1])
```

**Calculate the differences between the upper CIs and lower CIs and create a dataframe**
```{r}
df1$diff = df1$upper_ci - df1$lower_ci
df2$diff = df2$upper_ci - df2$lower_ci
df3 = data.frame(cbind(df1$diff, df2$diff))
head(df3)
```

**Calculate the difference between model 1 and model 2**
```{r}
df3$diff = df3$X1- df3$X2
```

```{r}
ggplot(df3, aes(diff)) + 
  geom_histogram(bins = 100) + 
  xlab("Diff") + 
  ylab("Count")
```

**The first model always has a larger difference between the upper CI and lower CI than the second model**


```{r}
display(fit_1)
```

**The between variance 11.65/(11.65 + 7.31) = 0.61446**

```{r}
display(fit_2)
```


**The between variance for the second model is 11.45/(11.45 + 7.32) = 0.6100, a little smaller than the between variance of the first model**

### (d) ###

**We can see that the second model is more precise because it shrinks the partial pooling closer to the complete pooling, so it's more precise**


## Problem 12.5 ##


**Create the group-level predictors log_uranium_full**
```{r}
library(rstanarm)

radon %>%
  mutate(log_uranium_full = log_uranium[county]) ->
  tidy_radon

head(tidy_radon)
```


**Fit the model using lmer with the group-level predictors and varying-intercept**
```{r}
fit_radon = lmer(log_radon ~ floor + log_uranium_full + (1| county), tidy_radon)
display(fit_radon)
```





